PCA

### 2019-11-10 analysis/11_pcalda.Rmd
new pca and tsne figures for talk

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
library(Rtsne)
library(viridis)
knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/')

```

@@

```{r pca}
# make pca
mypca <- prcomp(countData)

@@

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
head(mypcadf)

a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         shape = colData$tissue)) +
  geom_point() + labs(x = "PC1: 35.9%", y = "PC2: 32.5%") + theme_minimal(base_size = 6)  +
  theme(legend.title = element_blank(), legend.position = "left")
a

b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6))
b

ab <- plot_grid(a,b + theme(axis.text = element_blank()) + ggtitle(NULL), rel_widths = c(0.6,0.4))
ab
```

@@

# prep for adding columns
colData$V1 <- NULL
tsne_df_cols <- cbind(colData, tsne_df)

c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
  geom_point(size = 2) +
  theme_minimal(base_size = 6) +
  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
  theme(legend.position = "top") +
  labs(x = "tSNE 1", y = "tSNE 2")

d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = treatment)) +
  geom_point(size = 2) +
  theme_minimal(base_size = 6) +
  theme(legend.position = "top") +
  labs(x = "tSNE 1", y = "tSNE 2")

cd <- plot_grid(c,d)
cd
```

@@

```{r pca.tSNE, fig.width=6, fig.height=6}
plot_grid(ab,cd, ncol = 1)

plot_grid(a + theme(legend.position = "none"),
          b + theme(axis.text = element_blank()) + ggtitle(NULL), 
          c + theme(legend.position = "none"),
          d + theme(legend.position = "none"))
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
library(Rtsne)
library(viridis)

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/', cache = T)

# custom theme for plots
mytheme <- function(){
  theme_minimal(base_size = 12) + 
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "transparent", colour = NA),
          plot.background = element_rect(fill = "transparent", colour = NA),
          legend.title = element_blank())
}
```

@@

```{r}
# make pca dataframes
mypca <- prcomp(countData)

@@

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
head(mypcadf)
```

```{r pca}
a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         shape = colData$tissue,
                         alpha = colData$sex)) +
  geom_point(size = 2)  + theme_minimal(base_size = 6)  +
  mytheme() +
  theme(legend.title = element_blank(), 
        axis.text = element_blank()) +
  scale_alpha_manual(values = c(0.5,1)) +
  labs(x = "PC1: 35.9% of variance", y = "PC2: 32.5% of variance")

b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL)

png('../figures/pca/pca-1.png',width=600,height=600, units="px",bg = "transparent")
print(a)
dev.off()

png('../figures/pca/pca-2.png',width=600,height=600, units="px",bg = "transparent")
print(b)
dev.off()
```

```{r pcaloadings}
fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
fviz_contrib(mypca, choice = "var", axes = 2, top = 5) 
```

@@

# prep for adding columns
colData2 <- colData 
colData2$V1 <- NULL
tsne_df_cols <- cbind(colData2, tsne_df)

c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
  labs(x = "tSNE 1", y = "tSNE 2")

d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
  geom_point(size = 2) +
  mytheme() +
  labs(x = "tSNE 1", y = "tSNE 2") +
  scale_alpha_manual(values = c(0.5,1))

c
d

png('../figures/pca/tSNE-2.png',width=600,height=600, units="px",bg = "transparent")
print(c)
dev.off()

png('../figures/pca/tSNE-3.png',width=600,height=600, units="px",bg = "transparent")
print(d)
dev.off()
```

@@
